Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-04-03 WSSSPE2_report/WSSSPE2_review.tex
Sections 1 and 4: touch up.

\begin{abstract}      
This technical report records and discusses the Second Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE2). The workshop used an
alternative submission and peer-review process, which led to a set of papers
divided across five topic areas: exploring sustainability; software development
experiences; credit \& incentives; reproducibility \& reuse \& sharing; and code
testing \& code review. The report includes a description of the submission and
review process, the workshop keynote presentations, a set of lightning talks, a discussion on
sustainability, and the five discussions from the topic areas. The report also
contains a description of potential actions that were proposed in the the five
discussions.
\end{abstract}

@@

This report extends a previous report that discussed the submission,
peer-review, and peer-grouping processes in detail~\cite{WSSSPE2-pre-report}.
It is also based on a collaborative set of notes taken through a google doc
during the workshop~\cite{WSSSPE2-google-notes}. Overall, the report discusses
the organization work done before the workshop (\S\ref{sec:preworkshop}); the
keynotes (\S\ref{sec:keynotes}); a set of lightning talks (\S\ref{sec:lightning}),
intended to give an opportunity for attendees to quickly highlight an issue
or potential solution; a session on defining sustainability
(\S\ref{sec:defining}); five breakout sessions that explored sustainability
(\S\ref{sec:exploring}); software development experiences (\S\ref{sec:devel});
credit \& incentives (\S\ref{sec:credit}); reproducibility, reuse, \& sharing
(\S\ref{sec:reproduce}); code testing \& code review
(\S\ref{sec:code_testing}); and some conclusions (\S\ref{sec:conclusions}). The
report also includes an incomplete list of attendees (Appendix~\ref{sec:attendees}).

@@

Lightning talks were a new feature in WSSSPE2. Since the workshop program was
mostly dedicated to discussions, the organizers wanted to give the attendees a
chance to also `make a pitch' for an idea, either representing a contributed paper
or something different.  18 attendees volunteered to participate in the lightning
talks, each given two minutes to talk, and one slide if they wanted to use one.
The talked were presented in the reverse alphabetic order of speakers' last names.

@@

@@

@@

\item Alice Allen: Find it! Cite it!
\end{enumerate}

>>>>

\begin{abstract}      
This technical report records and discusses the Second Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE2). The workshop used an
alternative submission and peer-review process, which led to a set of papers
divided across five topic areas: exploring sustainability; software development
experiences; credit \& incentives; reproducibility \& reuse \& sharing; and code
testing \& code review. The report includes a description of the submission and
review process, the workshop keynote presentations, a series of lightning talks, a discussion on
sustainability, and the five discussions from the topic areas. The report also
contains a description of potential actions that were proposed in the the five
discussions.
\end{abstract}

@@

This report extends a previous report that discussed the submission,
peer-review, and peer-grouping processes in detail~\cite{WSSSPE2-pre-report}.
It is also based on a collaborative set of notes taken through a google doc
during the workshop~\cite{WSSSPE2-google-notes}. Overall, the report discusses
the organization work done before the workshop (\S\ref{sec:preworkshop}); the
keynotes (\S\ref{sec:keynotes}); a series of lightning talks (\S\ref{sec:lightning}),
intended to give an opportunity for attendees to quickly highlight an important issue
or a potential solution; a session on defining sustainability
(\S\ref{sec:defining}); five breakout sessions that explored sustainability
(\S\ref{sec:exploring}); software development experiences (\S\ref{sec:devel});
credit \& incentives (\S\ref{sec:credit}); reproducibility, reuse, \& sharing
(\S\ref{sec:reproduce}); code testing \& code review
(\S\ref{sec:code_testing}); and some conclusions (\S\ref{sec:conclusions}). The
report also includes an incomplete list of attendees (Appendix~\ref{sec:attendees}).

@@

\begin{comment}  
gvwilson@software-carpentry.org, Colin Venters <C.Venters@hud.ac.uk>,
j.spencer@imperial.ac.uk, erinrobinson@esipfed.org, marpierc@iu.edu,
jwpeterson@gmail.com, abani@buffalo.edu, clenhardt@renci.org, dsk@uchicago.edu,
Samin.Ishtiaq@microsoft.com, james@howison.name, s.a.harris@leeds.ac.uk,
marcus.hanwell@kitware.com, thabermann@hdfgroup.org, rdowns@ciesin.columbia.edu,
cboettig@gmail.com, jakob.blomer@cern.ch, editor@ascl.net 
\end{comment}

Lightning talks were a new feature in WSSSPE2. Since the workshop program was
mostly dedicated to discussions, the organizers wanted to give the attendees a
chance to also `make a pitch' for an idea, either representing a contributed paper
or something different.  Eighteen attendees volunteered to participate in the lightning
talks, each given only two minutes to speak and at most one slide to show.
The talks were presented in reverse alphabetic order of speakers' last names. 
In the rest of this session, we highlight the gist of the speakers' key messages.

@@

Howison sought to provoke discussion by proposing that papers whose workflows are
not kept current with the changing software ecosystem should be automatically
retracted. This would create an incentive for authors to keep their
software current and usable, rather than the current situation in which every
potential user has to do this individually. A softer version of the proposal
would identify papers whose software workflow has become bit-rotten and allow
others to keep the code up to date, either adding them as new authors of the
paper or providing credit in some other form for their academic service.

@@

Boettiger described how the rOpenSci project has been successful by focusing not just on
building software but also on building a community of researchers who learn and adopt
their approaches to reproducible research and sustainable software practice.
Through outreach, mentoring, workshops, and hackathons, they have not only reached
new users, but also turned users into co-developers of robust software and good practices to
support data science research across a growing set of disciplines.

@@

\item Alice Allen: Find it! Cite it!

The Astrophysics Source Code Library (ASCL) is an online registry of
scientist-written software used in astronomy research. Our primary interest is
rendering research more transparent by making this software more discoverable for
examination. The ASCL is treated like a publication by an indexing resource for
astronomy, the Astrophysics Data System (ADS). ADS tracks citations to what it
indexes, including citations to ASCL entries. Increasing rewards for writing
software, whether through citation, transitive credit or other methods, gives
software authors a powerful reason to take the time to build sustainability into
their software and is an excellent way to drive community change. 
\begin{comment}The ASCL
actively looks for opportunities to use change management strategies to move
astronomy to a more transparent discipline; these strategies can be used in
other disciplines to further change in them. 
\end{comment}
\end{enumerate}
